Latentcor: An R Package for Latent Correlation Estimation

### 2021-06-03 paper/paper.tex
Roughly fill up the paper.Rmd.

\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Latentcor: An R Package for Latent Correlation Estimation},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

Create a table

\begin{longtable}[]{@{}ll@{}}
\toprule
Type & continuous \\
\midrule
\endhead
continuous & (\textbf{davis2006reliable?}) \\
binary & (\textbf{dice1945measures?}) \\
truncated & (\textbf{dunne2002solutions?}) \\
ternary & (\textbf{bommert2020adjusted?}) \\
stabilityIntersectionGreedy & (\textbf{bommert2020adjusted?}) \\
stabilityIntersectionMBM & (\textbf{bommert2020adjusted?}) \\
stabilityIntersectionMean & (\textbf{bommert2020adjusted?}) \\
stabilityJaccard & (\textbf{jaccard1901etude?}) \\
stabilityKappa & (\textbf{carletta1996assessing?}) \\
stabilityLustgarten & (\textbf{lustgarten2009measuring?}) \\
stabilityNogueira & (\textbf{nogueira2018stability?}) \\
stabilityNovovicova & (\textbf{novovicova2009new?}) \\
stabilityOchiai & (\textbf{ochiai1957zoogeographic?}) \\
stabilityPhi & (\textbf{nogueira2016measuring?}) \\
stabilitySechidis & (\textbf{sechidis2020stability?}) \\
stabilitySomol & (\textbf{somol2008evaluating?}) \\
stabilityUnadjusted & (\textbf{bommert2020adjusted?}) \\
stabilityWald & (\textbf{wald2013stability?}) \\
stabilityYu & (\textbf{yu2012stable?}) \\
stabilityZucknick & (\textbf{zucknick2008comparing?}) \\
\bottomrule
\end{longtable}

type of variable (reference to papers) Definition for Kendall tau.
Describe the theorem states expected value of kendall tau = F(r). refer
to table for reference of formula. A sentence to describe two methods:
slow original vs interpolation. Table to show memory improvement compare
to mixedCCA. Generate data for pairs. One example of original call, one
for interpolation call.

@@

@@

>>>>

\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Latentcor: An R Package for Latent Correlation Estimation},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\begin{longtable}[]{@{}ll@{}}
\toprule
Type & continuous \\
\midrule
\endhead
continuous & Liu, Lafferty, \& Wasserman (2009) \\
binary & Fan, Liu, Ning, \& Zou (2017) \\
truncated & Yoon, Carroll, \& Gaynanova (2020) \\
ternary & Quan, Booth, \& Wells (2018) \\
\bottomrule
\end{longtable}

\emph{Definition 1} Fan et al.~(2017) considered the problem of
estimating \(\Sigma\) for the latent Gaussian copula model based on
Kendall's \(\tau\). Given the observed data
\((X_{1j}, X_{1k}), ..., (X_{nj}, X_{nk})\) for variables \(X_{j}\) and
\(X_{k}\), Kendall's \(\tau\) is defined as \[
\hat{\tau}_{jk}=\frac{2}{n(n-1)}\sum_{1\leq i <i'\leq n} sign(X_{ij}-X_{i'j})sign(X_{ik}-X_{i'k})
\] \emph{Theorem 1} Let \(W_{1}\in\cal{R}^{p_1}\),
\(W_{2}\in\cal{R}^{p_2}\), \(W_{3}\in\cal{R}^{p_3}\),
\(W_{4}\in\cal{R}^{p_4}\) be such that
\(W=(W_{1}, W_{2}, W_{3}, W_{4})\sim NPN(0,\Sigma,f)\) with
\(p=p_{1}+p_{2}+p_{3}+p_{4}\). Let
\(X=(X_{1}, X_{2}, X_{3}, X_{4})\in\cal{R}^{p}\) satisfy \(X_{j}=W_{j}\)
for \(j=1,...,p_{1}\), \(X_{j}=I(W_{j}>c_{j})\) for
\(j=p_{1}+1,...,p_{1}+p_{2}\), \(X_{j}=I(W_{j}>c_{j})W_{j}\) for
\(j=p_{1}+p_{2}+1,...,p\) and
\(X_{j}=I(W_{j}>c_{j}^{1})+I(W_{j}>c_{j}^{2})\) with
\(\Delta_{j}=f(c_{j})\), \(\Delta_{j}^{1}=f(c_{j}^{1})\) and
\(\Delta_{j}^{2}=f(c_{j}^{2})\). The rank-based estimator of \(\Sigma\)
based on the observed \(n\) realizations of \(X\) is the matrix
\(\hat{R}\) with \(\hat{r}_{jj}=1\),
\(\hat{r}_{jk}=\hat{r}_{kj}=F^{-1}(\hat{\tau}_{jk})\) with block
structure \[
\hat{R}=\left(\begin{array}\\
F^{-1}_{CC}(\hat{\tau})\hspace{.2in} F^{-1}_{CB}(\hat{\tau})\hspace{.2in} F^{-1}_{CT}(\hat{\tau})\hspace{.2in} F^{-1}_{CN}(\hat{\tau})\\
F^{-1}_{BC}(\hat{\tau})\hspace{.2in} F^{-1}_{BB}(\hat{\tau})\hspace{.2in} F^{-1}_{BT}(\hat{\tau})\hspace{.2in} F^{-1}_{BN}(\hat{\tau})\\
F^{-1}_{TC}(\hat{\tau})\hspace{.2in} F^{-1}_{TB}(\hat{\tau})\hspace{.2in} F^{-1}_{TT}(\hat{\tau})\hspace{.2in} F^{-1}_{TN}(\hat{\tau})\\
F^{-1}_{NC}(\hat{\tau})\hspace{.2in} F^{-1}_{NB}(\hat{\tau})\hspace{.2in} F^{-1}_{NT}(\hat{\tau})\hspace{.2in} F^{-1}_{NN}(\hat{\tau})
\end{array}\right)
\] The original method is taking estimated Kendall's \(\hat{\tau}\) and
other parameters to calculate latent correlation \(\hat{r}\). Whereas
the approximated method is using multilinear interpolation to
approximate latent correlation \(\hat{r}\) via pre-calculated grid
values (Yoon, Müller, \& Gaynanova, 2021).

refer to table for reference of formula.

Table to show memory improvement compare to mixedCCA.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(latentcor)}
\DocumentationTok{\#\#\# Data setting}
\NormalTok{n }\OtherTok{\textless{}{-}} \DecValTok{1000}\NormalTok{; p1 }\OtherTok{\textless{}{-}} \DecValTok{1}\NormalTok{; p2 }\OtherTok{\textless{}{-}} \DecValTok{1} \CommentTok{\# sample size and dimensions for two datasets.}
\NormalTok{maxcancor }\OtherTok{\textless{}{-}} \FloatTok{0.9} \CommentTok{\# true canonical correlation}

\DocumentationTok{\#\#\# Correlation structure within each data set}
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{0}\NormalTok{)}
\NormalTok{perm1 }\OtherTok{\textless{}{-}} \FunctionTok{sample}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\NormalTok{p1, }\AttributeTok{size =}\NormalTok{ p1);}
\NormalTok{Sigma1 }\OtherTok{\textless{}{-}} \FunctionTok{autocor}\NormalTok{(p1, }\FloatTok{0.7}\NormalTok{)[perm1, perm1]}
\NormalTok{blockind }\OtherTok{\textless{}{-}} \FunctionTok{sample}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{3}\NormalTok{, }\AttributeTok{size =}\NormalTok{ p2, }\AttributeTok{replace =} \ConstantTok{TRUE}\NormalTok{);}
\NormalTok{Sigma2 }\OtherTok{\textless{}{-}} \FunctionTok{blockcor}\NormalTok{(blockind, }\FloatTok{0.7}\NormalTok{)}
\NormalTok{mu }\OtherTok{\textless{}{-}} \FunctionTok{rbinom}\NormalTok{(p1}\SpecialCharTok{+}\NormalTok{p2, }\DecValTok{1}\NormalTok{, }\FloatTok{0.5}\NormalTok{)}

\DocumentationTok{\#\#\# true variable indices for each dataset}
\NormalTok{trueidx1 }\OtherTok{\textless{}{-}} \DecValTok{1}
\NormalTok{trueidx2 }\OtherTok{\textless{}{-}} \DecValTok{1}

\CommentTok{\# Data generation}
\NormalTok{simdata }\OtherTok{\textless{}{-}} \FunctionTok{GenerateData}\NormalTok{(}\AttributeTok{n=}\NormalTok{n, }\AttributeTok{trueidx1 =}\NormalTok{ trueidx1, }\AttributeTok{trueidx2 =}\NormalTok{ trueidx2,}
                        \AttributeTok{maxcancor =}\NormalTok{ maxcancor,}
                        \AttributeTok{Sigma1 =}\NormalTok{ Sigma1, }\AttributeTok{Sigma2 =}\NormalTok{ Sigma2,}
                        \AttributeTok{copula1 =} \StringTok{"exp"}\NormalTok{, }\AttributeTok{copula2 =} \StringTok{"cube"}\NormalTok{,}
                        \AttributeTok{muZ =}\NormalTok{ mu,}
                        \AttributeTok{type1 =} \StringTok{"binary"}\NormalTok{, }\AttributeTok{type2 =} \StringTok{"continuous"}\NormalTok{,}
                        \AttributeTok{c1 =} \FunctionTok{rep}\NormalTok{(}\DecValTok{1}\NormalTok{, p1), }\AttributeTok{c2 =}  \ConstantTok{NULL}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in GenerateData(n = n, trueidx1 = trueidx1, trueidx2 = trueidx2, : Same
## threshold is applied to the all variables in the first set.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{X1 }\OtherTok{\textless{}{-}}\NormalTok{ simdata}\SpecialCharTok{$}\NormalTok{X1; X2 }\OtherTok{\textless{}{-}}\NormalTok{ simdata}\SpecialCharTok{$}\NormalTok{X2; Sigma12\_tt }\OtherTok{\textless{}{-}}\NormalTok{ simdata}\SpecialCharTok{$}\NormalTok{Sigma12}
\CommentTok{\# Estimate latent correlation matrix with original method}
\NormalTok{R1\_tt\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X1, }\StringTok{"binary"}\NormalTok{, }\AttributeTok{method =} \StringTok{"original"}\NormalTok{)}
\NormalTok{R2\_tt\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X2, }\StringTok{"continuous"}\NormalTok{, }\AttributeTok{method =} \StringTok{"original"}\NormalTok{)}
\NormalTok{R12\_tt\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X1, }\AttributeTok{type1 =} \StringTok{"binary"}\NormalTok{, X2, }\AttributeTok{type2 =} \StringTok{"continuous"}\NormalTok{,}
                              \AttributeTok{method =} \StringTok{"original"}\NormalTok{)}\SpecialCharTok{$}\NormalTok{R12}
\CommentTok{\# Estimate latent correlation matrix with original method}
\NormalTok{R1\_tt\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X1, }\StringTok{"binary"}\NormalTok{, }\AttributeTok{method =} \StringTok{"approx"}\NormalTok{)}
\NormalTok{R2\_tt\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X2, }\StringTok{"continuous"}\NormalTok{, }\AttributeTok{method =} \StringTok{"approx"}\NormalTok{)}
\NormalTok{R12\_tt\_org }\OtherTok{\textless{}{-}} \FunctionTok{estR}\NormalTok{(X1, }\AttributeTok{type1 =} \StringTok{"binary"}\NormalTok{, X2, }\AttributeTok{type2 =} \StringTok{"continuous"}\NormalTok{,}
                              \AttributeTok{method =} \StringTok{"approx"}\NormalTok{)}\SpecialCharTok{$}\NormalTok{R12}
\end{Highlighting}
\end{Shaded}

@@

\leavevmode\hypertarget{ref-fan2017high}{}%
Fan, J., Liu, H., Ning, Y., \& Zou, H. (2017). High dimensional
semiparametric latent graphical model for mixed data. \emph{Journal of
the Royal Statistical Society. Series B: Statistical Methodology},
\emph{79}(2), 405--421.

\leavevmode\hypertarget{ref-liu2009nonparanormal}{}%
Liu, H., Lafferty, J., \& Wasserman, L. (2009). The nonparanormal:
Semiparametric estimation of high dimensional undirected graphs.
\emph{Journal of Machine Learning Research}, \emph{10}(10).

\leavevmode\hypertarget{ref-quan2018rank}{}%
Quan, X., Booth, J. G., \& Wells, M. T. (2018). Rank-based approach for
estimating correlations in mixed ordinal data. \emph{arXiv preprint
arXiv:1809.06255}.

@@

\leavevmode\hypertarget{ref-yoon2021fast}{}%
Yoon, G., Müller, C. L., \& Gaynanova, I. (2021). Fast computation of
latent correlations. \emph{Journal of Computational and Graphical
Statistics}, 1--8.
